Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2022-05-05 FDT2020/Introduction.tex
small edits

Carr and Vempala \cite{metarounding} showed if there an algorithm $\mathcal{A}$ that for $I\in \mathcal{I}$ and $c\geq 0$ finds $z\in S(I)$ with $cz\leq g'cx$ for any $x\in P(I)$, then $g'x$ can be written as a convex combination of points in $S(I)$. However, for many known problems there is a gap between the best known $g'$ and $g(\mathcal{I})$.

>>>>

Carr and Vempala \cite{metarounding} showed if there is an algorithm $\mathcal{A}$ that for $I\in \mathcal{I}$ and $c\geq 0$ finds $z\in S(I)$ with $cz\leq g'cx$ for any $x\in P(I)$, then $g'x$ can be written as a convex combination of points in $S(I)$. However, for many known problems there is a gap between the best known $g'$ and $g(\mathcal{I})$.

### 2022-05-05 FDT2020/main-section.tex
small edits

The FDT algorithm for 2EC iteratively applies Lemmas \ref{LPC2EC} and \ref{2ecpruning} to variables $x_1,\ldots,x_t$ to obtain leaf point solutions $L_t$. Finally, we just need to apply Lemma \ref{rounddown} to obtain the 2-edge-connected multi-subgraphs from every solution in $L_t$. Since $x$ is an extreme point we have $t\leq 2|V|-1$ \cite{boydpulley}. By Lemma \ref{2ecpruning} we have
\begin{align*}
\sum_{(x,\lambda)\in L_t} \frac{\lambda}{\sum_{(x,\lambda)\in L_t}\lambda} \floor{x} \leq \frac{1}{\sum_{(x,\lambda)\in L_t}\lambda} \sum_{(x,\lambda)\in L_t} \lambda {x} \leq g^t_{\2ec} x^*.
\end{align*}
\section{Computational Experiments with FDT}\label{sec:experiment}
We ran FDT on three network design problems: VC, TAP and 2EC. These experiments, especially those for 2EC, illustrate possible uses of FDT to study the integrality gap of a problem. We also compare FDT the original feasibility pump algorithm~\cite{fp1}\footnote{We coud not find code for feasibility pump 2.0~\cite{fp2}.} or problem-specific approximation algorithms. Our intent is to show that FDT's performance in practice is far better than the provable bound on integrality gap and the run time is reasonable. We are not recommending FDT as a replacement for other general heuristics, but it may have a place in a suite of heuristics for some problems.

>>>>

The FDT algorithm for 2EC iteratively applies Lemmas \ref{LPC2EC} and \ref{2ecpruning} to variables $x_1,\ldots,x_t$ to obtain leaf point solutions $L_t$. Finally, we just need to apply Lemma \ref{rounddown} to obtain the 2-edge-connected multi-subgraphs from every solution in $L_t$. Since $x$ is an extreme point we have $t\leq 2|V|-1$ \cite{boydpulley}. By Lemma \ref{2ecpruning} we have
\begin{align*}
\sum_{(x,\lambda)\in L_t} \frac{\lambda}{\sum_{(x,\lambda)\in L_t}\lambda} \floor{x} \leq \frac{1}{\sum_{(x,\lambda)\in L_t}\lambda} \sum_{(x,\lambda)\in L_t} \lambda {x} \leq g^t_{\2ec} x^*.
\end{align*}
\section{Computational Experiments with FDT}\label{sec:experiment}
We ran FDT on three network design problems: VC, TAP and 2EC. These experiments, especially those for 2EC, demonstrate using FDT to study the integrality gap of a problem. We also compare FDT to the original feasibility pump algorithm~\cite{fp1}\footnote{We could not find code for feasibility pump 2.0~\cite{fp2}.} or problem-specific approximation algorithms. Our intent is to show that FDT's performance in practice is far better than the provable bound on integrality gap and the run time is reasonable. We are not recommending FDT as a replacement for other general heuristics, but for some problems, it may have a place in a suite of heuristics.
